---
output: 
  github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
```



# mlmtools

Multilevel and mixed effects models often require specialized data pre-processing and further post-estimation derivations and graphics to gain insight into model results. `mlmtools` is a suite of pre- and post-estimation tools for multilevel models in R. The package's post-estimation tools are designed to work with models estimated using `lme4`'s lmer function, which fits linear mixed effects regression models. Although nearly all the functions provided in the `mlmtools` package exist as singleton functions within other R packages, they are often improved in `mlmtools` and more accessible by being located within a multilevel modeling specific package.

The package was developed by Jessica Mazen, Laura Jamison, Erik Ruzek, and Gus Sjobek.

## Included functions
* Intraclass correlation coefficients (ICC)
* Variance explained
  + When comparing nested models
  + $R^2$
* Test of necessity of random intercepts
* Visualizations 
  + Associations between variables within clusters
  + Associations between variables between clusters
  + Caterpillar plots


## Installation
To install the latest release version (0.0.0.9) from [GitHub](https://github.com/) with:

``` {r}
# install.packages("devtools")
devtools::install_github("lj5yn/mlmtools")
```

## Sample workflow

Working with the included data, we briefly show how some of the included functions can be used. 

```{r}
# data
library(mlmtools)
library(lme4)
data("instruction")

# fit variance components model
fit1 <- lmer(mathgain ~ 1 + (1|classid), instruction)
  
# intraclass correlation coefficient
ICCm(fit1)

# add predictor
fit2 <- lmer(mathgain ~ mathkind + (1|classid), instruction)

# variance explained by adding predictor
varCompare(fit2, fit1)

```

## Visualizations

Rich visualizations of associations can be had along with caterpillar plots, which graph the 95% prediction intervals for the random effects.

```{r}
# visualze within-group association
withinPlot(x = "mathkind", y = "mathgain", grouping = "classid", dataset = instruction)

# caterpillar plot
caterpillarPlot(fit2, grpvar = "classid")
```

